### 2021-12-14 thesis/tex/DCtheory.tex
update

The detector systematics include ice absorption and scattering, and overall, lateral, and head-on optical efficiencies of the DOMs. 
They are applied as correction factors using the best-fit points from the 2019 DeepCore $\nu_\tau$ appearance 
analysis~\cite{DC2019tauappearance}.

@@

\section{PINGU}
The methodology behind the PINGU simulations is the same as with our DeepCore study~. We use the public MC~\cite{PINGUdata}, which allows us to construct the event count as in Eq.~\ref{eq:MCevents}.
However, since no detector systematics is yet modelled for PINGU, the correction factors $C_{ijk}$ are all unity.
As with the DeepCore data, the PINGU Monte Carlo is divided into eight 
$\log_{10}E^{reco} \in [0.75,1.75]$ bins, and eight $\zreco \in [-1,1]$ bins for both track- and cascade-like events. 
We generate `data' by predicting the event rates at PINGU with the following best-fit parameters from~\cite{nufit}, except for the CP-violating phase which is set to zero for simplicity.

>>>>

The detector systematics include ice absorption and scattering and overall, lateral, and head-on optical efficiencies of the DOMs. 
They are applied as correction factors using the best-fit points from the 2019 DeepCore $\nu_\tau$ appearance 
analysis~\cite{DC2019tauappearance}.

@@

\section{PINGU}\label{ch:PINGUmethod}
The methodology behind the PINGU simulations is the same as with our DeepCore study~. We use the public MC~\cite{PINGUdata}, which allows us to construct the event count as in Eq.~\ref{eq:MCevents}.
However, since no detector systematics is yet modeled for PINGU, the correction factors $C_{ijk}$ are all unity.
As with the DeepCore data, the PINGU Monte Carlo is divided into eight 
$\log_{10}E^{reco} \in [0.75,1.75]$ bins, and eight $\zreco \in [-1,1]$ bins for both track- and cascade-like events. 
We generate `data' by predicting the event rates at PINGU with the following best-fit parameters from~\cite{nufit}, except for the CP-violating phase, which is set to zero for simplicity.

### 2021-12-14 thesis/tex/ICintro.tex
update

The flux is provided in \cite{hondaData,hondaArticle}, and a selection is shown in Table~\ref{table:flux}.
The flux data is binned in $\cos(\theta_z)$. The fluxes are averaged over the azimuthal direction and solar minimum/maximum. 
The units of the fluxes are given as \si{\per\GeV \per\metre\squared \per\second \per\steradian} and are omitted
from the table for clarity. 
We note that the fluxes for $\nt$ and $\bar{\nu}_{\tau}$ are missing. Kaons do not decay into $\nt$. Thus, we never have to use probabilities in the form 
$P_{\tau \beta}$, since we have no starting atmospheric $\nu_\tau$ flux. 
Interpolating the data yields makes us capable of approximating all four necessary fluxes for a given true energy and true zenith.
The result of $\phi_\mu$ is shown in Fig.~\ref{fig:flux}. We see that the flux is almost entirely zenith independent. 
% This is due to the 
% very small cross-section of the neutrinos. For particles interacting through the electromagnetic interaction, we would expect to see a 
% large difference between the flux of particles traversing through the whole diameter of the Earth, compared to particles that only penetrate a 
% small rim of the crust. Here, however, the flux at the South Pole is almost the same! This does not mean that the oscillation probability will be 
% equal. 
% First, the difference in the density profiles introduces wildly different matter effects. 
% Secondly, the zenith angle determines the baseline. Completely down-going neutrinos, with $\cos(\theta_z) = 0$ will be produced at an average distance of \SI{15}{\km}
% overhead, travel down to the icy sheet of Antarctica, and then penetrate the ice down \SI{2}{\km} to the DOMs~\cite{hondapaper}. 
The energy dependence of the flux is approximately $E^{-2.7}$, so our expected event count at higher energies will suffer from this strong decline in flux.

@@

In NC interactions, shown in Eq.~\ref{eq:NC}, no flavor change can occur, so the fermion leaves with increased energy.
The flavor and charge of the fermion or hadron $N$ interacting with the neutrino is unchanged, but energy 
from the neutrino will be transferred to the fermion that violently recoils, later producing a shower of tertiary particles.
\begin{align}\label{eq:NC}
    \nu_\alpha + N \to \nu_\alpha + N^\prime
\end{align}

In this work, we only study the detectors buried within the Antarctic ice by the IceCube collaboration. 
They are of Cherenkov type, which means that they detect 
a fermion by its emitted Cherenkov light, produced from its travel through the ice. 
If the secondary particles interact heavily with the ice, they will only travel a short distance before decaying 
into the shower which emit a localized flash of Cherenkov light. This event is referred to as a cascade.
NC interactions tend to be classified as cascades due to the production of a localized hadronic shower
However, CC $\ne$ interactions can also produce a similar localized flash of Cherenkov light, causing them to also contribute 
to the cascade events. So what about CC $\nt$? Due to the heaviness of the $\tau$, it rapidly decays - primarily into hadrons - giving us a short path
of travel with a final shower, thus manifesting themselves as a cascade~\cite{DC2019tauappearance}. A cascade event 
is shown in Fig.~\ref{fig:events_cascade} and Eq.~\ref{eq:cascade}, shows the possible interactions causing a cascade event.
Here $N$ denotes an incoming nucleon (consisting of one fermion due to interact with the neutrino), and the resultant $N^\prime$ can be a hadron or
a hadronic shower.
\begin{align}\label{eq:cascade}
    \nu_\alpha + N &\to \nu_\alpha + N^\prime \nonumber \\
    \nu_e + N &\to e + N^\prime  \nonumber \\
    \nu_\tau + N &\to \tau + N^\prime\,.
\end{align}

If the charged leptons don't interact as much in the ice, they penetrate a larger part of it, emitting light and tertiary particles
as they go. This event is referred to as a track and is often due to charged current muon interactions. A track event 
is shown in Fig.~\ref{fig:events_track}, and and Eq.~\ref{eq:track} shows the possible interactions causing a track event.
\begin{align}\label{eq:track}
    \nm + N \to \mu + N^\prime
\end{align}

@@

To detect the Cherenkov light, 60 Digital Optical Modules (DOMs) are placed on a long string up to \SI{17}{\metre} apart. 86 of these strings are then lowered
into \SI{2.5}{\km} deep boreholes in the ice. The holes are then sealed by refreezing the ice, resulting in a total of 5160 DOMs in a volume of approximately \SI{1}{\km^3}~\cite{weaverThesis}.

The strings and DOMs are not spaced evenly, making some parts of the detector more sensitive to certain energy ranges than others.
8 strings packed more tightly than the other 78, making that part of the detector sensitive to neutrino energies down to single-digit \si{\GeV}. This detector is referred to DeepCore. DeepCore will be treated as a separate and independent detector from the rest, which
retains the name IceCube. A view of the current setup can be seen in Fig.~\ref{fig:array}. In this work, we consider DeepCore data between \SIrange{5.6}{56}{\GeV} and IceCube data in the range \SIrange{0.5}{10}{\TeV}.
\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/icecube2.png}
    \caption{View of the full IceCube array, with the Eiffel tower for scale.}\label{fig:array}
\end{figure}

In 2017, the PINGU Letter of Intent was published~\cite{PINGUletter}. The `Precision IceCube Next Generation Upgrade' is an upgrade that will 
supplement DeepCore, i.e. boosting the capabilities of neutrino detection at the \si{\GeV} scale. As the PINGU upgrade is not yet financed nor built, we are
not able to use any data from it. However, the collaboration has released preliminary simulations which we will use to see how the upgrade might improve
results from IceCube and DeepCore. The PINGU simulations have the same structure as the DeepCore data, so our analysis referring to DeepCore will
also apply to PINGU except where noted. However, we treat the PINGU detector as independent of the other experiments.

\subsection{True and Reconstructed Parameters}
After an event has occurred, the IceCube collaboration \emph{reconstruct} the event. 
This means that, given the parameters recorded by the detector, what are their `true' values?
We are interested in two variables: energy and direction. Each event is tagged with a probable energy and zenith angle, 
called the reconstructed parameters $\Ereco$ and $\zreco$, which are the parameters according to the DOMs.
In other words, the collaboration uses the recorded data to backtrack the event, using the reconstructed parameters to approximate the true parameters. 
So a charged lepton hits the DOMs, and we ultimately end up with the associated neutrino's true energy and zenith angle, $\Etrue$ and $\ztrue$. 
The reconstructed parameters are what we are using to analyze the data (because this is what the detector sees), while the true parameters are used in the 
determination of that neutrino's `actual' flux and cross-section (because this is what Nature sees).

>>>>

The flux is provided in \cite{hondaData,hondaArticle}, and a selection is shown in Table~\ref{table:flux}.
The flux data is binned in $\cos(\theta_z)$. The fluxes are averaged over the azimuthal direction and solar minimum/maximum. 
The units of the fluxes are given as \si{\per\GeV \per\metre\squared \per\second \per\steradian} and are omitted
from the table for clarity. 
We note that the fluxes for $\nt$ and $\bar{\nu}_{\tau}$ are missing. Kaons do not decay into $\nt$. Thus, we never have to use probabilities in the form 
$P_{\tau \beta}$, since we have no starting atmospheric $\nu_\tau$ flux. 
Interpolating the data yields makes us capable of approximating all four necessary fluxes for a given true energy and true zenith.
The result of $\phi_\mu$ is shown in Fig.~\ref{fig:flux}. We see that the flux is almost entirely zenith independent. 
% This is due to the 
% very small cross-section of the neutrinos. For particles interacting through the electromagnetic interaction, we would expect to see a 
% large difference between the flux of particles traversing through the whole diameter of the Earth, compared to particles that only penetrate a 
% small rim of the crust. Here, however, the flux at the South Pole is almost the same! This does not mean that the oscillation probability will be 
% equal. 
% First, the difference in the density profiles introduces wildly different matter effects. 
% Secondly, the zenith angle determines the baseline. Completely down-going neutrinos, with $\cos(\theta_z) = 0$ will be produced at an average distance of \SI{15}{\km}
% overhead, travel down to the icy sheet of Antarctica, and then penetrate the ice down \SI{2}{\km} to the DOMs~\cite{hondapaper}. 
The energy dependence of the flux is approximately $E^{-2.7}$, so our expected event count at higher energies will suffer from this substantial decline in flux.

@@

In NC interactions, shown in Eq.~\ref{eq:NC}, no flavor change can occur, so the fermion leaves with increased energy.
The flavor and charge of the fermion or hadron $N$ interacting with the neutrino are unchanged. However, energy 
from the neutrino will be transferred to the fermion that violently recoils, later producing a shower of tertiary particles.
\begin{align}\label{eq:NC}
    \nu_\alpha + N \to \nu_\alpha + N^\prime
\end{align}

In this work, we only study the detectors buried within the Antarctic ice by the IceCube collaboration. 
They are of Cherenkov type, which means that they detect 
a fermion by its emitted Cherenkov light, produced from its travel through the ice. 
If the secondary particles interact heavily with the ice, they will only travel a short distance. They then decay
into a shower which emits a localized flash of Cherenkov light. This event is referred to as a cascade.
NC interactions tend to be classified as cascades due to the production of a localized hadronic shower
However, CC $\ne$ interactions can also produce a similar localized flash of Cherenkov light, causing them to also contribute 
to the cascade events. So what about CC $\nt$? Due to the heaviness of the $\tau$, it rapidly decays - primarily into hadrons - giving us a short path
of travel with a final shower, thus manifesting themselves as a cascade~\cite{DC2019tauappearance}. A cascade event 
is shown in Fig.~\ref{fig:events_cascade} and Eq.~\ref{eq:cascade}, shows the possible interactions causing a cascade event.
Here $N$ denotes an incoming nucleon (consisting of one fermion due to interact with the neutrino), and the resultant $N^\prime$ can be a hadron or
a hadronic shower.
\begin{align}\label{eq:cascade}
    \nu_\alpha + N &\to \nu_\alpha + N^\prime \nonumber \\
    \nu_e + N &\to e + N^\prime  \nonumber \\
    \nu_\tau + N &\to \tau + N^\prime\,.
\end{align}

If the charged leptons do not interact as much in the ice, they penetrate a larger part of it, emitting light and tertiary particles
as they go. This event is called a track and is often due to charged current muon interactions. A track event 
is shown in Fig.~\ref{fig:events_track}, and and Eq.~\ref{eq:track} shows the possible interactions causing a track event.
\begin{align}\label{eq:track}
    \nm + N \to \mu + N^\prime
\end{align}

@@

In order to detect the Cherenkov light, 60 Digital Optical Modules (DOMs) are placed on a long string up to \SI{17}{\metre} apart. 86 of these strings are then lowered
into \SI{2.5}{\km} deep boreholes in the ice. The holes are then sealed by refreezing the ice, resulting in a total of 5160 DOMs in a volume of approximately \SI{1}{\km^3}~\cite{weaverThesis}.

The strings and DOMs are not spaced evenly, making some parts of the detector more sensitive to certain energy ranges than others.
8 strings packed more tightly than the other 78, making that part of the detector sensitive to neutrino energies down to single-digit \si{\GeV}. This detector is referred to DeepCore. 
DeepCore will be treated as a separate and independent detector from the rest of the array, which
retains the name IceCube. A view of the current setup can be seen in Fig.~\ref{fig:array}. In this work, we consider DeepCore data between \SIrange{5.6}{56}{\GeV} and IceCube data in the range \SIrange{0.5}{10}{\TeV}.
\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/icecube2.png}
    \caption{View of the full IceCube array, with the Eiffel tower for scale.}\label{fig:array}
\end{figure}

In 2017, the PINGU Letter of Intent was published~\cite{PINGUletter}. The `Precision IceCube Next Generation Upgrade' is an upgrade that will 
supplement DeepCore, i.e. boosting the capabilities of neutrino detection at the \si{\GeV} scale. As the PINGU upgrade is not yet financed nor built, we are
unable to use any data from it. However, the collaboration has released preliminary simulations, which we will use. The simulated events 
are between \SIrange{5.6}{56}{\GeV}, the same range as the DeepCore data.
The PINGU simulations have the same structure as the DeepCore data, so our analysis referring to DeepCore will
also apply to PINGU except where noted. However, we treat the PINGU detector as independent of the other experiments.

\subsection{True and Reconstructed Parameters}
After an event has occurred, the IceCube collaboration \emph{reconstruct} the event. 
Given the parameters recorded by the detector, what are their `true' values?
We are interested in two variables: energy and direction. Each event is tagged with a probable energy and zenith angle, 
called the reconstructed parameters $\Ereco$ and $\zreco$, which are the parameters according to the DOMs.
In other words, the collaboration uses the recorded data to backtrack the event, using the reconstructed parameters to approximate the true parameters. 
So a charged lepton hits the DOMs, and we ultimately end up with the associated neutrino's true energy and zenith angle, $\Etrue$ and $\ztrue$. 
The reconstructed parameters are what we are using to analyze the data (because this is what the detector sees), while the true parameters are used in the 
determination of that neutrino's `actual' flux and cross-section (because this is what Nature sees).

### 2021-12-14 thesis/tex/ICtheory.tex
update

% \documentclass{thesis}
% \usepackage[margin=0.5in]{geometry}
% \input{commands}
% \begin{document}

\section{IceCube}\label{ch:ICmethod}
As the neutrinos have propagated the Earth, they arrive at the South Pole, where they interact with charged leptons in the ice. The charged lepton then emits the Cherenkov light that is detected by the array.
We construct the event rate for each bin as
\begin{align}\label{eq:ICevents}
    N_{ij} &= T \sum_\beta\int_{(\cos{\theta_z^r})_i}^{(\cos{\theta_z^r})_{i+1}} \dd \cos{\theta^r_z} \int_{E^r_{j}}^{E^r_{j+1}} \dd E^r 
    \int_0^\pi R(\theta^r,\theta^t) \dd \cos{\theta^t} \int_0^\infty R(E^r,E^t) \phi_\beta^\text{det}  A^\text{eff}_\beta \dd E^t
    \,,
\end{align}
where $T$ is the live time of the detector, $\beta$ the final neutrino flavors, $\theta_z^r$ the reconstructed 
zenith angle, i.e. the deduced direction of the incoming neutrino binned with index $i$. $\theta^t_z$ is the true zenith angle, i.e. the actual direction of the incoming neutrino. 
$E^r$ is the reconstructed energy, binned with index $j$. $R(\theta^r,\theta^t)$ is a zenith resolution function 
that describes the relationship with the reconstructed and true zenith angles, specific to the 86-string configuration of IceCube.
$R(E^r,E^t)$ is an energy resolution function 
that describes the relationship with the reconstructed and true energies. $\phi_\beta^\text{det}$ is the conventional atmospheric neutrino flux for flavor $\beta$, propagated to detector level
in accordance with Eq.~\ref{eq:propFlux}.

We now are interested in the effective area $\Aeff$, 
i.e.~the cross-sectional area of the detector that the lepton is exposed to.
$\Aeff$ depends on several parameters, some of them being detector physical volume, $\Etrue$, $\ztrue$, and the particle type. 
Fortunately, the binned $\Aeff$ is provided to us by the collaboration~\cite{ICaeff}.
An excerpt from this data is shown in Table~\ref{table:aeff}.

@@

Here, $\Aeff$ has been averaged over $\Aeff_\mu$ and $\Aeff_{\bar{\mu}}$ by the collaboration. Thus, both $\mu$ and $\bar\mu$ will on average, experience the same $\Aeff$ in our model. 
Just as with the fluxes, we interpolate this in $\Etrue$ and $\ztrue$, and show the result in Fig.~\ref{fig:aeff}
Since the IceCube array is slightly rectangular, the zenith angle affects the cross-sectional area to which the array the leptons are exposed to.
While the flux was almost flat in $\ztrue$, the introduction of the zenith dependent $\Aeff$ will make the result slightly more zenith dependent than the flux itself. Energy-wise, the situation here is completely reverse from the flux. 
Increasing linearly with energy, the effective area of the detector approaches its geometrical area of \SI{1e6}{\metre^2} but is still only in the single-digit range at \si{\TeV} energies.

@@

So now we have the physical quantities in the true parameters. 
But as we discussed, we need a way to translate this into the reconstructed parameters that the detector gives us. We will call the relationship between 
$\Ereco$ and $\Etrue$ the energy resolution function, and the relationship between $\zreco$ and $\ztrue$ the zenith resolution function. 
We assume the relationship to follow a logarithmic Gaussian distribution, giving it the form 
\begin{align}\label{eq:gaussian}
    R(x^r, x^t) = \frac{1}{\sqrt{2\pi} \sigma_{x^r}x^r} \exp\left[-\frac{(\log x^r-\mu(x^t))^2}{2\sigma_{x^r}^2}\right]\,.
\end{align}
The parameters of the Gaussian are $\sigma_{x^r}(x^t)$ and $\mu(x^t)$, which are functions of the true parameters. By multiplying the Gaussian in Eq.~\ref{eq:gaussian}, we are reweighing the values by the 
probability density of that point. This process is also called \emph{smearing} because it effectively spreads out the data around a certain point.

@@

First, we let $\zreco = \ztrue$ for all values. The angular resolution in IceCube for track-like events is less than $\SI{2}{\degree}$, making $\ztrue$ coincide with $\zreco$ for our study~\cite{IC2020}.
Thus, we only need to concern ourselves with the energy resolution.
In Fig.~\ref{fig:IC_MC_gpr}, we have plotted all event counts found in the MC file, over 8 million. However, this is too much data to process efficiently, with many outliers that ultimately don't weigh in 
that much in the final event count. To resolve this, we have opted to train a Gaussian process regressor on the dataset, from which we can extract the predicted mean and standard deviation for a point.
When doing this over $\Ereco$, we sample $\Etrue$ in the 99th percentile around the predicted mean. We then obtain the shaded band shown in Fig.~\ref{fig:IC_MC_gpr}. %TODO add plot with Etrue vs prob dens

@@

Now, Eq.~\ref{eq:ICevents} handles the Gaussian smearing, but we are not provided systematic error sources, DOM efficiencies, and other nuisance parameters. To correct this,
we will aim to come as close as possible to the IceCube Monte Carlo, and then normalize with it. That way, we know that our null hypotheses will align while we are free to form additional hypotheses with different 
physics parameters. %The binned Monte Carlo events that IceCube used as their null hypothesis in the 2020 sterile analysis is shown in %TODO:MC vs myMC fig.

@@

\begin{figure}[ht]
    \centering
    \includegraphics[scale=0.4]{figures/IC_data.pdf}
    \caption{IceCube track events from~\cite{IC2020}.}\label{fig:IC_data}
\end{figure}

\subsection*{Monte Carlo normalization}
Independent researchers outside of the IceCube collaboration will not be able to more precisely
simulate the detector. The IceCube Monte Carlo is a complex and proprietary machinery, so our goal in this 
section is to come as close as we can to their Monte Carlo simulations. After we are confident that 
our code displays the same overall features as the `official', we normalize our results $N_{ij}^\text{sim}$ as 
\begin{align}\label{eq:MC_norm}
    N_{ij} = \frac{N_{ij}^\text{null}}{N_{ij}^\text{MC}} N_{ij}^\text{sim}\,.
\end{align}
For each bin $i,j$, we then obtain a correction factor that contains information that we are unable
to obtain or sufficiently incorporate. One example of such information is the systematic errors of the DOMs.
Recent IceCube data releases do not include such information. Since the systematic errors are affecting the 
event count on a bin-by-bin basis, they can in theory drastically modify the binned results. Another example of
an error source that will be remedied by this method is the flux. We are using a fairly simple model of the atmospheric 
flux that excludes atmospheric prompt and astrophysical fluxes. The IceCube collaboration uses several different flux models which are initialized 
by a parametrization of the cosmic ray flux.\footnote{Included in the cosmic ray models are e.g. the pion to kaon 
ratio, which are often used as a nuisance parameter. By not being able to include this in our error analysis, our method will 
be limited to only consider the overall flux normalization, rather than the components that produce the flux in the first place.}

>>>>

\section{IceCube}\label{ch:ICmethod}
As the neutrinos have propagated the Earth, they arrive at the South Pole, where they interact with charged leptons in the ice. The charged lepton then emits the Cherenkov light detected by the array.
We construct the event rate for each bin as
\begin{align}\label{eq:ICevents}
    N_{ij} &= T \sum_\beta\int_{(\cos{\theta_z^r})_i}^{(\cos{\theta_z^r})_{i+1}} \dd \cos{\theta^r_z} \int_{E^r_{j}}^{E^r_{j+1}} \dd E^r 
    \int_0^\pi R(\theta^r,\theta^t) \dd \cos{\theta^t} \int_0^\infty R(E^r,E^t) \phi_\beta^\text{det}  A^\text{eff}_\beta \dd E^t
    \,,
\end{align}
where $T$ is the live time of the detector, $\beta$ the final neutrino flavors, $\theta_z^r$ the reconstructed 
zenith angle, i.e.~the deduced direction of the incoming neutrino binned with index $i$. $\theta^t_z$ is the true zenith angle, i.e.~the actual direction of the incoming neutrino. 
$E^r$ is the reconstructed energy, binned with index $j$. $R(\theta^r,\theta^t)$ is a zenith resolution function 
that describes the relationship with the reconstructed and true zenith angles, specific to the 86-string configuration of IceCube.
$R(E^r,E^t)$ is an energy resolution function 
that describes the relationship with the reconstructed and true energies. $\phi_\beta^\text{det}$ is the conventional atmospheric neutrino flux for flavor $\beta$, propagated to detector level
in accordance with Eq.~\ref{eq:propFlux}.

We now are interested in the effective area $\Aeff$, 
i.e.~the cross-sectional area of the detector that the lepton is exposed to.
$\Aeff$ depends on several parameters, some of them being physical detector volume, $\Etrue$, $\ztrue$, and the particle type. 
Fortunately, the binned $\Aeff$ is provided to us by the collaboration~\cite{ICaeff}.
An excerpt from this data is shown in Table~\ref{table:aeff}.

@@

Here, $\Aeff$ has been averaged over $\Aeff_\mu$ and $\Aeff_{\bar{\mu}}$ by the collaboration. Thus, both $\mu$ and $\bar\mu$ will, on average, experience the same $\Aeff$ in our model. 
Just as with the fluxes, we interpolate this in $\Etrue$ and $\ztrue$, and show the result in Fig.~\ref{fig:aeff}
Since the IceCube array is slightly rectangular, the zenith angle affects the cross-sectional area to which the array the leptons are exposed to.
While the flux was almost flat in $\ztrue$, the introduction of the zenith dependent $\Aeff$ will make the result slightly more zenith dependent than the flux itself. 
Increasing linearly with energy, the effective area of the detector approaches its geometrical area of \SI{1e6}{\metre^2} but is still only in the single-digit range at \si{\TeV} energies.

@@

So now we have the physical quantities in the true parameters. 
As we discussed, we need a way to translate this into the reconstructed parameters that the detector gives us. We will call the relationship between 
$\Ereco$ and $\Etrue$ the energy resolution function, and the relationship between $\zreco$ and $\ztrue$ the zenith resolution function. 
We assume the relationship to follow a logarithmic Gaussian distribution, giving it the form 
\begin{align}\label{eq:gaussian}
    R(x^r, x^t) = \frac{1}{\sqrt{2\pi} \sigma_{x^r}x^r} \exp\left[-\frac{(\log x^r-\mu(x^t))^2}{2\sigma_{x^r}^2}\right]\,.
\end{align}
The parameters of the Gaussian are $\sigma_{x^r}(x^t)$ and $\mu(x^t)$, which are functions of the true parameters. By multiplying the Gaussian in Eq.~\ref{eq:gaussian}, we are reweighing the values by the 
probability density of that point. This process is also called \emph{smearing} because it effectively spreads out the data around a certain point.

@@

First, we let $\zreco = \ztrue$ for all values. The angular resolution in IceCube for track-like events is less than $\SI{2}{\degree}$, making $\ztrue$ coincide with $\zreco$ for our study~\cite{IC2020}.
Thus, we only need to concern ourselves with the energy resolution function.
In Fig.~\ref{fig:IC_MC_gpr}, we have plotted all event counts found in the MC file, over 8 million. However, this is too much data to process efficiently, with many outliers that ultimately do not weigh in 
that much in the final event count. To resolve this, we have opted to train a Gaussian process regressor on the dataset, from which we can extract the predicted mean and standard deviation for a point.
When doing this over $\Ereco$, we sample $\Etrue$ in the 99th percentile around the predicted mean. We then obtain the shaded band shown in Fig.~\ref{fig:IC_MC_gpr}.

@@

Now, Eq.~\ref{eq:ICevents} handles the Gaussian smearing, but we are not provided systematic error sources, DOM efficiencies, and other nuisance parameters. To correct this,
we will aim to come as close as possible to the IceCube Monte Carlo, and then normalize with it. That way, we know that our null hypotheses will align while we are free to form additional hypotheses with different 
physics parameters.

@@

\begin{figure}[ht]
    \centering
    \includegraphics[scale=0.4]{figures/IC_data.pdf}
    \caption{IceCube track events from~\cite{IC2020}, displaying the $E^-{-1.7}$ drop-off in event count.}\label{fig:IC_data}
\end{figure}

\subsection*{Monte Carlo normalization}
Independent researchers outside of the IceCube collaboration will not be able to simulate the detector more precisely.
The IceCube Monte Carlo is a complex and proprietary machinery, so our goal in this 
section is merely to come as close as we can to it. After we are confident that 
our code displays the same overall features as the `official', we normalize our results $N_{ij}^\text{sim}$ as 
\begin{align}\label{eq:MC_norm}
    N_{ij} = \frac{N_{ij}^\text{null}}{N_{ij}^\text{MC}} N_{ij}^\text{sim}\,.
\end{align}
For each bin $i,j$, we then obtain a correction factor that contains information that we are unable
to obtain or sufficiently incorporate. One example of such information is the systematic errors of the DOMs.
Recent IceCube data releases do not include such information. Since the systematic errors are affecting the 
event count on a bin-by-bin basis, they can, in theory, drastically modify the binned results. Another example of
an error source that will be remedied by this method is the flux. We are using a fairly simple model of the atmospheric 
flux that excludes atmospheric prompt and astrophysical fluxes. The IceCube collaboration uses several different flux models, which are initialized 
by a parametrization of the cosmic ray flux.\footnote{Included in the cosmic ray models are e.g. the pion to kaon 
ratio, which are often used as a nuisance parameter. By not being able to include this in our error analysis, our method will 
be limited to only consider the overall flux normalization rather than the components that produce the flux in the first place.}

### 2021-12-14 thesis/tex/SM.tex
update

% \input{commands}\begin{document}
\begin{figure}[!h]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/SM_figure.pdf}
\end{figure}
\section{The Standard Model}\label{ch:SM}
In order to describe the three quantizable forces of nature, we gather the mediators of each force -- the vector bosons -- into local (gauge) symmetry groups. 
Each vector boson has one corresponding generator, the set of which constitutes the group.
The strong charge is mediated by eight massless gluons, which correspond to the eight independent generators of $\text{SU}(3)_C$. 
The weak charge is mediated by the three massive gauge bosons $W^\pm$ and $Z$, and the massless photon $\gamma$, 
which constitute the generators of $\text{SU}(2)_L$ and $\text{U}(1)_Y$.

@@

\section{Mass Generation}
In the Standard Model, fermion masses are generated by the Higgs mechanism through Yukawa couplings with the fermion's right and left-handed components.
All neutrinos are left-handed and all antineutrinos are right-handed~\cite{giunti}. This observational fact is reflected in the 
left-handed lepton doublet and right-handed lepton singlet

@@

We extend the Standard Model by adding a right-handed component of this field with neutrino Yukawa couplings $Y_{\alpha \beta}^{\prime \nu}$ to the Higgs-lepton Yukawa part of the interaction Lagrangian 
\begin{align} 
    \mathcal{L}_{int} \subseteq -\left( \frac{v + H}{\sqrt{2}} \right) \left[ \bar{\ell}^{\prime}_{\alpha L} Y_{\alpha \beta}^{\prime \ell} \ell_{\beta R}^{\prime} + \bar{\nu}_{\alpha L}^{\prime} Y_{\alpha \beta}^{\prime \nu} \nu_{\beta R}^{\prime}\right]\,,
\end{align}
where $v$ is the Higgs vacuum expectation value, $H$ the Higgs field, $\ell_{\alpha L}^\prime (\ell_{\beta R}^\prime)$ the left (right) handed charged lepton field, and summation over the lepton indices $\alpha,\beta$ is implied. $Y_{\alpha \beta}^{\prime \ell}$ are the charged lepton Yukawa couplings, and $\nu^\prime_{\alpha L}(\nu^\prime_{\beta R})$ are the left (right) handed neutrino fields .
Now, the Yukawa coupling matrices $Y^{\prime \ell}$ and $Y^{\prime \nu}$ are non-diagonal, as emphasized with their prime. 
This can be remedied by diagonalizing $Y^{\prime \ell}$  with a biunitary transformation
\begin{align}
    V_{ L}^{\ell \dagger} Y^{\prime \ell} V_{ R}^{\ell}=Y^{\ell} \,.
\end{align}
where $V^\ell_L$ and $V^\ell_R$ are two unitary matrices.
Similarly, we diagonalize and $Y^{\prime \nu}$ with a unitary matrix $V^\nu$. We note that this will not yet generate neutrino masses, 
since no flavor rotation can make massless fields massive.
Now we state a crucial difference between the properties of the charged lepton and the neutrino fields.
While the charged lepton flavor eigenstate was uniquely determined by its mass eigenstate, the neutrino flavor is a superposition of mass eigenstates. 
This is because neutrinos are indirectly detected via the observation of its associated charged lepton, so there is no requirement for neutrino flavor eigenstates to have a definite mass. 
The flavor of a neutrino is then, by definition, the flavor of the associated charged lepton. 
This fact is denoted as giving the mass eigenstates Latin numerals and letters, while the flavor eigenstates stay as Greek letters.

>>>>

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/SM_figure.pdf}
\end{figure}
\section{The Standard Model}\label{ch:SM}
In order to describe the three quantizable forces of nature, we gather the mediators of each force -- the vector bosons -- into local (gauge) symmetry groups. 
Each vector boson has one corresponding generator, the set of which constitutes the group.
The strong charge is mediated by eight massless gluons, which correspond to the eight independent generators of $\text{SU}(3)_C$. 
The weak charge is mediated by the three massive gauge bosons $W^\pm$ and $Z$, and the massless photon $\gamma$, 
which constitute the generators of $\text{SU}(2)_L$ and $\text{U}(1)_Y$.

@@

\section{Mass Generation}
In the Standard Model, fermion masses are generated by the Higgs mechanism through Yukawa couplings with the fermion's right and left-handed components.
All neutrinos are left-handed, and all antineutrinos are right-handed~\cite{giunti}. This observational fact is reflected in the 
left-handed lepton doublet and right-handed lepton singlet

@@

We extend the Standard Model by adding a right-handed component of this field with neutrino Yukawa couplings $Y_{\alpha \beta}^{\prime \nu}$ to the Higgs-lepton Yukawa part of the interaction Lagrangian 
\begin{align} 
    \mathcal{L}_{int} \subseteq -\left( \frac{v + H}{\sqrt{2}} \right) \left[ \bar{\ell}^{\prime}_{\alpha L} Y_{\alpha \beta}^{\prime \ell} \ell_{\beta R}^{\prime} + \bar{\nu}_{\alpha L}^{\prime} Y_{\alpha \beta}^{\prime \nu} \nu_{\beta R}^{\prime}\right]\,,
\end{align}
where $v$ is the Higgs vacuum expectation value, $H$ the Higgs field, $\ell_{\alpha L}^\prime (\ell_{\beta R}^\prime)$ the left (right) handed charged lepton field, and summation over the lepton indices $\alpha,\beta$ is implied. $Y_{\alpha \beta}^{\prime \ell}$ are the charged lepton Yukawa couplings, and $\nu^\prime_{\alpha L}(\nu^\prime_{\beta R})$ are the left (right) handed neutrino fields .
Now, the Yukawa coupling matrices $Y^{\prime \ell}$ and $Y^{\prime \nu}$ are non-diagonal, as emphasized with their prime. 
This can be remedied by diagonalizing $Y^{\prime \ell}$  with a biunitary transformation
\begin{align}
    V_{ L}^{\ell \dagger} Y^{\prime \ell} V_{ R}^{\ell}=Y^{\ell} \,.
\end{align}
where $V^\ell_L$ and $V^\ell_R$ are two unitary matrices.
Similarly, we diagonalize and $Y^{\prime \nu}$ with a unitary matrix $V^\nu$. 
Now we state a crucial difference between the properties of the charged lepton and the neutrino fields.
While the charged lepton flavor eigenstate was uniquely determined by its mass eigenstate, the neutrino flavor is a superposition of mass eigenstates. 
This is because neutrinos are indirectly detected via the observation of its associated charged lepton, so there is no requirement for neutrino flavor eigenstates to have a definite mass. 
The flavor of a neutrino is then -- by definition -- the flavor of the associated charged lepton. 
This fact is denoted as giving the mass eigenstates Latin numerals and letters, while the flavor eigenstates stay as Greek letters.
